---
title: "Exam 3 Summary"
format:
  html:
    theme: flatly
    toc: false
    number-sections: true
    css: styles.css
<<<<<<< HEAD
    code-fold: true
=======
>>>>>>> 035259fc5832dbad87d2f32b907a27c5c26101af
---

## Overview

This summary reviews the main ideas from Exam 3: Loops I and II, API I and II, Web Scraping, Databases and SQL, and Interactive Visualization.

## Loops I + II

### What it is

Loops are used to repeat code. Control structures are used to control the flow of code based on conditions. In R, common control structures include `if`, `else`, `for`, `while`, `repeat`, `break`, and `next`.

### Why it matters

Loops help automate repeated tasks. Instead of writing the same code many times, I can let R go through values one by one. This is useful when I need to process multiple elements in a vector, list, or data frame.

### Key ideas / key functions

- `if`
- `else`
- `for`
- `while`
- `break`
- `next`
- `seq_along()`
- writing functions
- `map()` from `purrr` as another way to iterate

### Detailed explanation

A `for` loop is the most common loop in R. It takes a variable and gives it values from a sequence one at a time. Then it runs the code inside the braces for each value. This is useful when I want to repeat the same action across multiple items.

`if` and `else` help R make decisions. If a condition is true, one block of code runs. If it is false, another block can run instead. This is important when the action depends on the data.

A big idea from this topic is that repeated code should often become a function. If I copy and paste code many times, it is usually better to turn it into a function with arguments. This reduces mistakes, improves readability, and makes the code easier to update.

Another related idea is iteration with `purrr`. The `map()` family does repeated operations without writing a full loop. For example, `map()` returns a list, `map_dbl()` returns a numeric vector, and `map_chr()` returns a character vector.

### Simple example
```{r}
x <- c(2, 4, 6)

for (i in seq_along(x)) {
  print(x[i] * 2)
}
```

### What to remember

Loops help me repeat code, but if I keep repeating the same pattern, I should think about writing a function or using map().


## API I + II

### What it is

An API is a way for one program to communicate with another program and request data. Many APIs return data in JSON format. JSON stands for JavaScript Object Notation, and it is a lightweight text format that stores data using name/value pairs and ordered lists of values. This makes it common in web-based services and REST APIs.

### Why it matters

APIs let me bring data directly into R from online sources. This matters because it is usually cleaner, faster, and more reliable than copying data manually from a webpage. In R, the `jsonlite` package is especially useful because it can convert JSON into regular R objects without needing a lot of manual restructuring. That means API data can become vectors, data frames, or matrices that I can work with right away.

### Key ideas / key functions

- API
- REST API
- JSON
- `fromJSON()`
- `toJSON()`
- simplification
- vectors
- data frames
- matrices
- pagination

### Detailed explanation

The `jsonlite` package is designed to translate JSON data into R objects and back again. One important idea is **simplification**. When JSON is read into R with `fromJSON()`, arrays of primitives can become atomic vectors, arrays of objects can become data frames, and arrays of equal-length arrays can become matrices. This is helpful because JSON often comes in nested structures, and simplification makes the result easier to analyze.

Another important point is that API data is often live data. Instead of reading a saved file, I can call a URL directly and let R request the current data. The materials show examples of this with GitHub, CitiBike, and Ergast. In practice, this means I can use one function call to bring online data into R.

Some APIs also return large amounts of data across multiple pages. In that case, I may need to make repeated requests, store each response, and then combine them. This connects API work back to loops and iteration, because I may need a `for` loop or a `map()` workflow to collect everything.


### What to remember

If data is available through an API, that is usually the cleanest way to get it into R because the data is already structured for machines to read.


## Web Scraping and URLs

### What it is

Web scraping means collecting data from webpage content. A URL is the address of a web resource, and it tells the browser or R where to find something online. Understanding URLs is part of understanding how online data is organized.

### Why it matters

Before collecting data from the web, I need to know what kind of source I am dealing with. Sometimes the source is a webpage meant for humans to read. Other times it is an API meant for programs to read. If I understand the structure of URLs, I can better figure out where the data is, how the site is organized, and what changes when query parameters are added.

### Key ideas / key functions

- web scraping
- URL
- scheme
- host
- path
- parameters
- fragment
- relative path
- absolute path
- URL encoding
- `url_parse()`
- `param_get()`
- `url_encode()`
- `url_decode()`

### Detailed explanation

A URL usually has several parts. The **scheme** tells the protocol, such as `http` or `https`. The **host** is the domain name. The **path** tells the location of the specific resource. Some URLs also include **query parameters**, which pass extra information to the server, and **fragments**, which point to a specific part of a page.

This matters because many websites and APIs change their output depending on the query parameters. For example, a URL might contain something like `?pageid=1023&export=json`. Those values are extra instructions that tell the server what to return. The materials explain that URLs can also include credentials, a port number, and fragment identifiers, depending on the use case.

The `urltools` package is useful for breaking apart and inspecting URLs. It can parse a URL into parts, extract parameter values, add new parameters, or remove old ones. This is helpful when I need to understand a complicated web address instead of just copying it blindly.

This topic also connects to relative paths in R projects and Quarto files. A relative path depends on the current file location, while an absolute path gives the full location. That idea matters for local files, while full URLs matter for online resources. Both are really about understanding location and structure.

### Simple example

```{r}
library(urltools)

url <- "http://en.wikipedia.org/wiki/api.php?action=parse&pageid=1023&export=json"

url_parse(url)
param_get(url, parameter_names = c("pageid", "export"))
```

### What to remember

Before collecting web data, I should first decide whether I am looking at a webpage to scrape or an API that already gives structured data, because that changes the best method to use.

## Databases and SQL

### What it is

A database stores data in tables. SQL, which stands for Structured Query Language, is the main language used to query and manage those tables. In R, I can connect to a database and work with its tables instead of treating everything like a normal in-memory data frame.

### Why it matters

Databases are useful when data is too large, too complex, or too frequently updated to keep reloading as regular files. Instead of loading everything into R, I can connect to the database and ask for only the specific data I need. This is more efficient and more realistic for large-scale data work.

### Key ideas / key functions

- database
- table
- SQL
- DBMS
- `dbConnect()`
- `dbWriteTable()`
- `dbReadTable()`
- `dbListTables()`
- `dbGetQuery()`
- `tbl()`
- `show_query()`
- `collect()`
- lazy evaluation
- `SELECT`
- `FROM`
- `WHERE`
- `GROUP BY`
- `ORDER BY`

### Detailed explanation

A database can be thought of as a collection of tables. These tables are similar to data frames, but they are stored differently and are designed for large-scale storage and retrieval. The materials explain that database tables usually live on disk, can be very large, and often have indexes that help find rows quickly. Databases are managed by DBMS systems, and these can be client-server systems, cloud systems, or in-process systems like duckdb.

In R, the `DBI` package provides the low-level interface for connecting to databases and running commands. The `dbplyr` package builds on this by letting me use familiar `dplyr` verbs while translating them into SQL behind the scenes.

A very important idea here is **lazy evaluation**. When I work with a database table through `dbplyr`, R often does not run the full query immediately. Instead, it builds up the query step by step. This is helpful because it avoids pulling huge datasets into R too early. I can use `show_query()` to see the SQL translation, and I use `collect()` only when I want the final result brought into R.

SQL itself is based on key clauses. `SELECT` chooses columns, `FROM` says which table to use, `WHERE` filters rows, `GROUP BY` creates groups for summaries, and `ORDER BY` sorts the output. Even if I mostly use `dplyr`, I still need to understand these SQL ideas because that is what is actually happening underneath.

### Simple example

```{r}
library(DBI)
library(dbplyr)
library(dplyr)

con <- dbConnect(duckdb::duckdb())
dbWriteTable(con, "mpg", ggplot2::mpg)

mpg_db <- tbl(con, "mpg")

mpg_db |>
  filter(class == "suv") |>
  select(manufacturer, model, hwy) |>
  show_query()
```

### What to remember

Databases help me work with large data efficiently, and dbplyr lets me keep using familiar dplyr syntax while SQL does the work underneath.

## Interactive Visualization

### What it is

Visualization is the process of showing data in a way that helps people understand it. A good plot is not only correct but also clear, focused, and easy to read. In this topic, the main idea is that a plot should communicate a message, not just display information.

### Why it matters

In data analysis, making the plot is only part of the job. I also need to make sure the audience can understand the plot quickly. This matters because the people reading the graph may not know the data well, and they may not know the variable names or the context. Good labels, good titles, annotations, and better design choices make the graph much more effective.

### Key ideas / key functions

- `ggplot()`
- `aes()`
- `labs()`
- `geom_text()`
- `geom_label()`
- `geom_label_repel()`
- `geom_text_repel()`
- `theme()`
- `annotate()`
- `geom_hline()`
- `geom_vline()`
- title
- subtitle
- caption
- annotations
- meaningful labels
- data-to-ink ratio

### Detailed explanation

A key idea in this topic is that an exploratory plot and a communication plot are not the same thing. An exploratory plot is often quick and simple because it is only for my own thinking. A communication plot needs more effort because it should help someone else understand the main message.

The `labs()` function is one of the easiest ways to improve a plot. I can use it to create stronger titles, subtitles, captions, and clearer axis labels. A good title should summarize the main finding, not just describe the graph type.

Annotations are also important. Functions like `geom_text()` and `geom_label()` let me label individual observations or groups. When labels overlap too much, `geom_label_repel()` and `geom_text_repel()` help spread them out so the graph stays readable.

Another major idea is reducing clutter. The materials discuss things like meaningful labels, useful references, and a higher data-to-ink ratio. This means that the ink in the plot should mostly be used for important information instead of unnecessary decoration. A cleaner theme, fewer distracting grid elements, and more direct labeling usually make a plot stronger.

This topic also connects to design choices like legends, reference lines, and captions. If used well, these help guide the reader’s attention and make the message of the graph more obvious.

### What to remember

A strong plot should make the main point clear quickly. Clear labels, useful annotations, and less clutter usually make the visualization much better.


