---
title: "Homework 01 Brazilian Companies Capital Stock Analysis"
---


## TidyTuesday Section
Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-01-27/readme.md}).  Provide sufficient background for readers to grasp your narrative.

## Data Exploration

```{r}
install.packages("tidytuesdayR")
tuesdata <- tidytuesdayR::tt_load('2026-01-27')
tuesdata <- tidytuesdayR::tt_load(2026, week = 4)

companies <- tuesdata$companies
legal_nature <- tuesdata$legal_nature
qualifications <- tuesdata$qualifications
size <- tuesdata$size

companies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-01-27/companies.csv')
legal_nature <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-01-27/legal_nature.csv')
qualifications <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-01-27/qualifications.csv')
size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-01-27/size.csv')
```

```{r}
head(companies)
```

## Research Question
Within each company size category, which type of owner qualifications are most common in these firms have the highest capital?

## Data Cleaning
```{r}
library(dplyr)
library(ggplot2)

top10 <- companies %>%
  group_by(company_size) %>%
  filter(capital_stock >= quantile(capital_stock, 0.90, na.rm = TRUE)) %>%
  ungroup()

top10_counts <- top10 %>%
  count(company_size, owner_qualification, name = "n") %>%
  group_by(company_size) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup()

top10_top5 <- top10_counts %>%
  group_by(company_size) %>%
  slice_max(order_by = pct, n = 5, with_ties = FALSE) %>%
  ungroup()

```

```{r}
head(top10_top5)
summary(top10_top5$pct)
```


## Visualization
```{r}
ggplot(top10_top5, aes(x = pct, y = owner_qualification)) +
  geom_col() +
  facet_wrap(~ company_size, scales = "free_y") +
  labs(
    title = "Most common owner qualifications among the highest-capital firms",
    subtitle = "Highest-capital = top 10% of capital_stock within each company size group",
    x = "Share of firms in top 10% (within size)",
    y = "Owner qualification"
  ) +
  theme_minimal()

```














