---
title: "Homework 04"
format: 
  html: 
    embed-resources: true
---


## TidyTuesday Section (optional)

::: {.callout-important title="Instructions"}

You can count work on this week's TidyTuesday toward the exceptional work required for an A in the Homework component.

:::


Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}).  Provide sufficient background for readers to grasp your narrative.

#### Read Data

```{r}
library(tidyverse)
library(tidytuesdayR)
library(lubridate)
library(scales)

sfi_grants <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-02-24/sfi_grants.csv')
glimpse(sfi_grants)
```

### Research Question

Is SFI funding concentrated in a few institutions? (2000 - 2024)


### Data Wrangling

```{r}
sfi_clean <- sfi_grants |>
  mutate(year = year(start_date),
         funding = current_total_commitment,
         research_body = str_squish(research_body)) |>
  filter(!is.na(year), !is.na(research_body), !is.na(funding)) |>
  filter(year >= 2000, year <= 2024)
```

```{r}
nrow(sfi_clean)
summary(sfi_clean$funding)
count(sfi_clean, year) |> arrange(year)
```

Top 10 Institutions 
```{r}
top10_share <- sfi_clean |>
  group_by(research_body) |>
  summarise(total_funding = sum(funding), .groups = "drop") |>
  mutate(share = total_funding / sum(total_funding)) |>
  arrange(desc(total_funding)) |>
  slice_head(n = 10)

top10_total_share <- sum(top10_share$share)
top10_total_share
```


```{r}
range(top5_share$top5_share)
```

### Data Visualizations

```{r}
ggplot(top10_share, aes(x = reorder(research_body, share), y = share)) +
  geom_col(fill = "#2C7BB6") +
  geom_text(
    aes(label = scales::dollar(total_funding, prefix = "€", scale = 1e-6, suffix = "M")),
    hjust = -0.05, size = 3.5
  ) +
  coord_flip() +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 0.1),
    expand = expansion(mult = c(0, 0.12))  # adds room for labels
  ) +
  labs(
    title = "Top 10 Institutions by Share of Total SFI Funding (2000–2024)",
    subtitle = paste0("Combined, the top 10 institutions received ",
                      scales::percent(top10_total_share, accuracy = 0.1),
                      " of all SFI funding"),
    x = NULL,
    y = "Share of total funding",
    caption = "Source: Science Foundation Ireland Grants Data (TidyTuesday 2026 Week 8). Author: Jerry Miao"
  ) +
  theme_minimal(base_size = 13)
```

