---
title: "Homework 06 - TidyTuesday (Pi Digits) "
format: 
  html: 
    embed-resources: true
---

## Dataset Import From TidyTuesday
```{r}
library(tidytuesdayR)
library(tidyverse)

tuesdata <- tidytuesdayR::tt_load(2026, week = 12)
pi_digits <- tuesdata$pi_digits
```


## Research Question
Are the digits of π evenly distributed across 0–9, and does the distribution become more uniform as we include more digits?

## Data Exploration
```{r}
glimpse(pi_digits)
```
```{r}
digit_freq <- pi_digits %>%
  count(digit) %>%
  mutate(prop = n / sum(n))

digit_freq
```

## Visualizations

#### Visualization 1: Overall Digit Distribution
```{r}
ggplot(digit_freq, aes(x = factor(digit), y = prop)) +
  geom_col(fill = "steelblue") +
  geom_hline(yintercept = 0.1, linetype = "dashed", color = "red") +
  labs(
    title = "Digits of π distribution",
    x = "Digit",
    y = "Proportion",
    caption = "Red dashed line represents perfect uniform distribution (10%)"
  ) +
  theme_minimal()
```

All digits appear with roughly equal proportions near 10%, suggesting that π’s digits are uniformly distributed.

#### Visualization 2: Does Distribution Stabilize Over Time?
```{r}
pi_cumulative <- pi_digits %>%
  mutate(index = row_number()) %>%
  group_by(digit) %>%
  mutate(cum_count = row_number()) %>%
  ungroup() %>%
  group_by(digit) %>%
  mutate(cum_prop = cum_count / index) %>%
  ungroup()

ggplot(pi_cumulative, aes(x = index, y = cum_prop, color = factor(digit))) +
  geom_line(alpha = 0.7) +
  geom_hline(yintercept = 0.1, linetype = "dashed") +
  labs(
    title = "Digit frequencies converge toward 10%",
    x = "Digit Position",
    y = "Cumulative Proportion",
    color = "Digit"
  ) +
  theme_minimal()
```
Although proportions fluctuate early, they quickly stabilize near 10%, demonstrating convergence toward a uniform distribution.


#### Visualization 3: Runs of Repeated Digits
```{r}
pi_runs <- pi_digits %>%
  mutate(run = cumsum(digit != lag(digit, default = first(digit)))) %>%
  group_by(run, digit) %>%
  summarize(length = n(), .groups = "drop")

ggplot(pi_runs, aes(x = length)) +
  geom_histogram(binwidth = 1, fill = "darkorange") +
  labs(
    title = "Most repeated digit runs are short",
    subtitle = "Long runs exist but are rare",
    x = "Run Length",
    y = "Count"
  ) +
  theme_minimal()
```

```{r}
pi_cumulative %>%
  filter(index <= 50000) %>%  
  ggplot(aes(x = index, y = cum_prop, color = factor(digit))) +
  geom_line(linewidth = 1, alpha = 0.8) +
  geom_hline(yintercept = 0.1, linetype = "dashed", color = "black") +
  labs(
    title = "Randomness emerges quickly in the digits of π",
    subtitle = "Each line represents one digit (0–9). The dashed line indicates 10%.",
    caption = "Source: TidyTuesday (2026-03-24) One Million Digits of Pi Author: Jerry Miao.",
    x = "Digit Position (First 50,000 digits)",
    y = "Cumulative Proportion",
    color = "Digit"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot",
    legend.position = "right"
  )
```

## Conclusion
Although the distribution of digits in π appears uneven at the beginning, this is largely due to small sample size effects. As more digits are observed, the proportions of each digit quickly stabilize around 10%. This pattern demonstrates a clear convergence toward a uniform distribution, suggesting that the digits of π behave like a random sequence in the long run. Overall, the visualization highlights how apparent randomness emerges as the dataset grows.
