---
title: "Homework 06 TidyTuesday (Pi Digits) "
execute:
  echo: true
  warning: false
  message: false
format:
  html:
    embed-resources: false
---

## Dataset Import From TidyTuesday
```{r}
library(tidytuesdayR)
library(tidyverse)

tuesdata <- tidytuesdayR::tt_load(2026, week = 12)
pi_digits <- tuesdata$pi_digits
```


## Research Question
Are the digits of π evenly distributed across 0–9, and does the distribution become more uniform as we include more digits?

## Data Exploration
```{r}
glimpse(pi_digits)
```
```{r}
digit_freq <- pi_digits %>%
  count(digit) %>%
  mutate(prop = n / sum(n))

digit_freq
```

## Visualizations

#### Visualization 1: Overall Digit Distribution
```{r}
ggplot(digit_freq, aes(x = factor(digit), y = prop)) +
  geom_col(fill = "steelblue") +
  geom_hline(yintercept = 0.1, linetype = "dashed", color = "red") +
  labs(
    title = "Digits of π distribution",
    x = "Digit",
    y = "Proportion",
    caption = "Red dashed line represents perfect uniform distribution (10%)"
  ) +
  theme_minimal()
```

All digits appear with roughly equal proportions near 10%, suggesting that π’s digits are uniformly distributed.

#### Visualization 2: Does Distribution Stabilize Over Time?
```{r}
pi_cumulative <- pi_digits %>%
  mutate(index = row_number()) %>%
  group_by(digit) %>%
  mutate(cum_count = row_number()) %>%
  ungroup() %>%
  group_by(digit) %>%
  mutate(cum_prop = cum_count / index) %>%
  ungroup()

ggplot(pi_cumulative, aes(x = index, y = cum_prop, color = factor(digit))) +
  geom_line(alpha = 0.7) +
  geom_hline(yintercept = 0.1, linetype = "dashed") +
  labs(
    title = "Digit frequencies converge toward 10%",
    x = "Digit Position",
    y = "Cumulative Proportion",
    color = "Digit"
  ) +
  theme_minimal()
```
Although proportions fluctuate early, they quickly stabilize near 10%, demonstrating convergence toward a uniform distribution.


#### Visualization 3: Runs of Repeated Digits
```{r}
pi_cumulative_clean <- pi_digits %>%
  mutate(
    index = digit_position,
    digit = factor(digit)
  ) %>%
  group_by(digit) %>%
  mutate(cum_count = row_number()) %>%
  ungroup() %>%
  mutate(cum_prop = cum_count / index) %>%
  filter(index >= 1000, index <= 50000) %>%
  mutate(index_bin = floor(index / 1000) * 1000) %>%
  group_by(digit, index_bin) %>%
  summarize(cum_prop = last(cum_prop), .groups = "drop")

pi_end_labels <- pi_cumulative_clean %>%
  group_by(digit) %>%
  filter(index_bin == max(index_bin)) %>%
  ungroup()
```

```{r}
library(ggrepel)
library(scales)

pi_end_labels <- pi_cumulative_clean %>%
  group_by(digit) %>%
  filter(index_bin == max(index_bin)) %>%
  ungroup()

ggplot(pi_cumulative_clean, aes(x = index_bin, y = cum_prop, color = digit)) +
  geom_hline(
    yintercept = 0.10,
    linetype = "dashed",
    linewidth = 0.7,
    color = "gray40"
  ) +
  geom_rect(
    aes(xmin = 0, xmax = 5000, ymin = -Inf, ymax = Inf),
    inherit.aes = FALSE,
    fill = "gray90",
    alpha = 0.35
  ) +
  geom_line(linewidth = 0.9, alpha = 0.85) +
  geom_text_repel(
    data = pi_end_labels,
    aes(label = digit),
    nudge_x = 2500,
    direction = "y",
    hjust = 0,
    segment.color = NA,
    size = 3.5,
    show.legend = FALSE
  ) +
  annotate(
    "text",
    x = 46000,
    y = 0.103,
    label = "Expected share = 10%",
    size = 3.5,
    color = "gray25",
    hjust = 1
  ) +
  annotate(
    "text",
    x = 2500,
    y = 0.128,
    label = "Early positions\nfluctuate more",
    size = 3.3,
    color = "gray35"
  ) +
  scale_x_continuous(
  labels = scales::number_format(scale = 1 / 1000, suffix = "k"),
  breaks = seq(0, 50000, 10000),
  limits = c(0, 54000)
  ) +
  scale_y_continuous(
  labels = scales::percent_format(accuracy = 1),
  limits = c(0.07, 0.13),
  breaks = seq(0.07, 0.13, 0.01)
  ) +
  labs(
    title = "The first 50,000 digits of π quickly settle near 10%",
    subtitle = "Digit frequencies fluctuate early, then converge toward the expected uniform share.",
    x = "Digit position",
    y = "Cumulative share of each digit",
    caption = "Source: One Million Digits of Pi dataset (TidyTuesday 2026-03-24) | Author: Jerry Miao | Date: 2026-03-24"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 17),
    plot.subtitle = element_text(size = 11),
    plot.caption = element_text(size = 8, hjust = 0),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "gray88")
  )
```

## Conclusion
Although the distribution of digits in π appears uneven at the beginning, this is largely due to small sample size effects. As more digits are observed, the proportions of each digit quickly stabilize around 10%. This pattern demonstrates a clear convergence toward a uniform distribution, suggesting that the digits of π behave like a random sequence in the long run. Overall, the visualization highlights how apparent randomness emerges as the dataset grows.
